Psych 254 W15 PS #3

### 2015-03-21 analyses/ps3.Rmd
Started on ps3
Worked on ps3

@@

@@

```{r}
double.sample <- function () {
}
```

@@

@@

@@

@@

@@

@@

@@

@@

Thought question: Take a moment to interpret the coefficients of the model. 
Q1 - What are the units?
Q2 - How does the interaction relate to the plot?
Q3 - Should there be an interaction in the model? What does it mean? How important is it?

Now make predictions from the model you like the best. What should happen with
doses of 0, 1.5, 2.5, and 10 under both supplements?

@@

@@

>>>>

```{r}
sigCount = 0
numSamples = 10000
for(i in 1:numSamples) {
  fakeSample = rnorm(30)
  stat = t.test(fakeSample)
  if(stat$p.value < 0.05) {
    sigCount = sigCount + 1;
  }
}
print(sigCount / numSamples)
```

@@

```{r}
numSamples = 10000
testRes <- replicate(numSamples, t.test(rnorm(30))$p.value < .05)
cat("false positive rate:", sum(testRes)/numSamples)
```

**Using both methods, we find a false positive rate of about 0.05! In other words, if the null hypothesis is true and there is no effect, we'll still report one about 5% of the time. This is our alpha value (in decision-theoretic terms). As the number of samples increases, we'll converge to this rate exactly. **

@@

```{r}
double.sample <- function (maxP) {
  fakeSample1 = rnorm(30)
  test1p = t.test(fakeSample1)$p.value
  if(test1p >= maxP || test1p < .05) {
    return (test1p < 0.05)
  } else {
    fakeSample2 = rnorm(30)
    test2p = t.test(c(fakeSample1, fakeSample2))$p.value
    return (test2p < 0.05)
  }
}
```

@@

```{r}
numSamples = 10000
sum(replicate(numSamples, double.sample(.25)))/numSamples
```

@@

**There is indeed an inflation of false positives: we find false positives about 7% of the times instead of 5% as before**

@@

```{r}
numSamples = 10000
d <- data.frame(caps = c(.05, .25, .5, .75, 1)) %>%
  group_by(caps) %>%
  mutate(falseAlarm = sum(replicate(numSamples, double.sample(caps)))/numSamples)
d
```

@@

**Inflation continues to grow up to about 9% in the worse-case (when you always run 30 more trials after observing $p>0.05$). This means you will be about twice as likely to incorrectly report a significant finding when no effect exists. This is pretty bad. **

@@

```{r}
data(ToothGrowth)
qplot(y = len, x = dose, color = supp, data = ToothGrowth)
```

@@

```{r}
qplot(len, data = ToothGrowth)
```

** We see that length has a rather wide distribution, possibly multimodal. It is likely a mixture of different distributions, which we can investigate in our linear model.**

@@

```{r}
mainEffectModel <- lm(len ~ dose + supp, data = ToothGrowth)
summary(mainEffectModel)
interactionModel <- lm(len ~ dose * supp, data = ToothGrowth)
summary(interactionModel)
anova(mainEffectModel, interactionModel, test="Chisq")
```

**We see that both dose and supplement are significant in the main effects model, and that the interaction between them is significant in the interactive model.**

@@

```{r}
mainEffectModel <- lm(len ~ -1 + dose + supp, data = ToothGrowth)
summary(mainEffectModel)
interactionModel <- lm(len ~ -1 + dose * supp, data = ToothGrowth)
summary(interactionModel)
```

**This did not affect the interaction coefficient, but it added a level of the supplement factor**

Thought question: Take a moment to interpret the coefficients of the model. 
Q1 - What are the units?
**The coefficient in front of dose is the average slope of the line if the two supplement groups were collapsed together. The coefficients in front of the dose groups are the numbers added to this baseline dose at each level to get their respective group means.**
Q2 - How does the interaction relate to the plot?
**In the plot, we saw that the slope across doses is steeper for VC supplement than OJ supplement. This is the interaction our model is picking up on.** 
Q3 - Should there be an interaction in the model? What does it mean? How important is it?
**Yes, a nested model comparison shows that the interactive model accounts for a significantly higher proportion of the variance than the additive model, $\chi^2(1) =  88.92, p = 0.02$.**
Now make predictions from the model you like the best. What should happen with
doses of 0, 1.5, 2.5, and 10 under both supplements?

@@

```{r}
newD = data.frame(dose=rep(c(0,1.5,2.5,10),2), supp = c(rep('OJ', 4), rep('VC', 4)))
newD
newD$predictions = predict(interactionModel, newD)
newD
ggplot(newD, aes(x = dose, y = predictions, color = supp)) + geom_line()
```

@@

```{r}
plot(resid(interactionModel))
```

**Not bad! No clear trends, but also quite a bit of variability**
